---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data
Sorry if my english is not good. 
Unzip the file in your working directory and load with R:
```{r}
library("lubridate")
library("ggplot2")
suppressMessages(library("dplyr"))
library("tidyr")
data<-read.csv("activity.csv")
data$date<-ymd(data$date)

steps_day<-aggregate(steps~date,data,sum,na.rm=TRUE)

```


## What is mean total number of steps taken per day?

making a histogram for steps per day

```{r}
summary(steps_day$steps)
rango<-diff(range(steps_day$steps))
mean_step<-mean(steps_day$steps)
median_step<-median(steps_day$steps)

ggplot(steps_day, aes(x=steps), fill="green")+ geom_histogram(binwidth=rango/15, fill="green", colour="black")+theme_bw()+geom_vline(xintercept=mean_step, colour="red", size=1)+ggtitle("Steps per Day")

```

- The mean of steps per day is **`r format(mean_step, digits=2)`**
- The median of the steps per day is: **`r format(median_step, digits=2)`**


## What is the average daily activity pattern?

  We need agrupapate the data by interval and generate the line plot
  
  
```{r}
steps_interval<-aggregate(steps~interval,data,mean,na.rm=TRUE)

ggplot(steps_interval, aes(x = interval, y = steps)) + geom_line(color = "blue")+ theme_bw()+ ggtitle("Average number of steps by interval")

interval_of_maxsteps<-steps_interval[which.max(steps_interval$steps), ]$interval
```

  The interval where the mean of steps is max: **`r interval_of_maxsteps`**

## Imputing missing values

  We replace the NA values for the mean (easy way for imputing values)
  
```{r}
count_missing<-sum(is.na(data))
```

  There are **`r count_missing`** missing values.
  
  
```{r}
imput_data<-data

for (i in 1:nrow(data)) {
    if (is.na(data$steps[i])) {
        imput_data$steps[i] = steps_interval[steps_interval$interval == data$interval[i],]$steps
    }
}

```

  Compare de imput data y the original data
  
```{r}
head(data)#original data
head(imput_data) #data with mean of steps by interval
imput_steps_day<-aggregate(steps~date,imput_data,sum)
```

  Lets calculate de mean, median and do the histogram
  
```{r}

imput_rango<-diff(range(imput_steps_day$steps))
mean_imput_step<-mean(imput_steps_day$steps)
median_imput_step<-median(imput_steps_day$steps)

ggplot(imput_steps_day, aes(x=steps), fill="green")+ geom_histogram(binwidth=imput_rango/15, fill="green", colour="black")+theme_bw()+geom_vline(xintercept=mean_imput_step, colour="red", size=1)+ggtitle("Steps per Day in the imput data")
```

- The mean of steps per day is **`r format(mean_imput_step, digits=2)`**
- The median of the steps per day is: **`r format(median_imput_step, digits=2)`**

  The median now is much clouser to the mean.
  
```{r}
imput_interval_of_maxsteps<-imput_data[imput_data$steps == max(imput_data$steps), ]$interval
```

 and the interval with thw max number of steps is **`r imput_interval_of_maxsteps`**
   
## Are there differences in activity patterns between weekdays and weekends?

  Firts, we need to create a varible "dia", that contain if the date is a weekend or weekday. 

```{r}
# LC_TIME="Spanish_Spain.1252" (Sys.getlocale("LC_TIME"))
# In my region Weekdays are Lunes, Martes, Miercoles, Jueves y Viernes and
# weekends are Sabado y Domingo


imput_data$dia<-ifelse(weekdays(imput_data$date)=="domingo" |weekdays(imput_data$date)=="sábado",c("weekend"),c("weekday"))

imput_data$dia<-as.factor(imput_data$dia)
```
  
  Now, plot the line graph for weekdays and weekends
  
```{r}
library(lattice)
imput_steps_dia <- aggregate(steps ~ interval + dia, imput_data, mean)
xyplot(steps ~ interval | dia, data = imput_steps_dia, layout = c(1, 2), type = "l")
```
